A Comparison of Graph Processing Systems

### 2020-08-29 documentation/paper/sections/overview/giraph.tex

Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop\cite{Giraph}. Hadoop as a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.
Because of the underlying Hadoop MapReduce infrastructure, expanding single-node processing to a multi-node cluster is almost seamless.

Giraph is based on computation units that communicate using messages and are synchonized with barriers.

@@

Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex again.
A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep is reached the computation ends.

Each vertex outputs some local information (e.g. the final vertex value) as result.

Giraph being an Apache project makes it the most actively maintained and tested project in our comparison. Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{https://gitbox.apache.org/repos/asf?p=giraph.git}.

>>>>

Apache Giraph is an iterative graph processing framework, built on top of Apache Hadoop \cite{Giraph}. Hadoop is a large MapReduce infrastructure providing a reliable (fault tolerant) basis for large scale graph processing.

TODO // MapReduce hier oder wo anders erklären?
MapReduce is a in three phases
\begin{enumerate}
	\item Map phase: The input data is distributed between a set of map processes, the map functionality is specified by the user. Ideally all map processes run in parallel so the map processes need to be independent. Results from this phase are written into (multiple) intermediate storage points.
	\item Shuffle phase: The results are grouped according to a key provided by the map algorithm. Each set of results is then handed to one system for the next phase.
	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The reduce functionality is again specified by the user and ideally runs in parallel.
\end{enumerate}
Giraph uses the Map functionality of Hadoop to run the algorithms, Reduce is only used as an identity function.

Giraph's computation model is closely related to the BSP model discussed in \autoref{sec:bsp}. This means that Giraph is based on computation units that communicate using messages and are synchonized with barriers.

@@

Any vertex can stop computing (i.e. setting its state to inactive) at any time but incoming messages will reactivate the vertex again.
A vote-to-halt method is applied, i.e. if all vertices are inactive or if a user defined superstep is reached the computation ends.
Once calculation is finished, each vertex outputs some local information (e.g. the final vertex value) as result.

Because of the underlying Hadoop infrastructure, expanding single-node processing to a multi-node cluster is seamless. Hadoop supplies a distributed file system (HDFS), on which all computation is performed. Giraph is thus, even when only using a single node, running in a distributed manner.

Giraph being an Apache project makes it the most actively maintained and tested project in our comparison. Over the course of this project, several new updates were pushed to Giraph's source repository\footnote{\url{https://gitbox.apache.org/repos/asf?p=giraph.git}}.

### 2020-08-29 documentation/paper/sections/results/sssp.tex

\subsection{Single-source Shortest-paths}

\subsubsection{Single-node}
\autoref{fig:singleNodeSSSP} shows the average calculation times (time without initialization overhead), execution time and the normalized overhead for SSSP on the different frameworks. In these figures, Galois with 96 threads is shown. We show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
\begin{figure*}[ht!]
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
		\caption{Calculation times for SSSP on a single node}
		\label{fig:singleNodeSSSP_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
		\caption{Execution times for SSSP on a single node}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
		\caption{Overhead time normalized by the graph size in million edges}
		\label{fig:singleNodeSSSP_overheadNormalized}
	\end{subfigure}
	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.
	The runs on rMat28 for Ligra and Polymer failed and the frameworks were unable to complete the task.}
	\label{fig:singleNodeSSSP}
\end{figure*}

\subsubsection{Distributed}
\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
		\caption{Calculation times for distributed SSSP}
		\label{fig:distributedSSSP_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
		\caption{Execution times for distributed SSSP}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
	\label{fig:distributedSSSP}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
	\label{fig:distributedSSSP_overhead}
\end{figure}

For the distributed scenario, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.

Results are especially interesting for Giraph since it seems to not cope well with synthetic graphs. Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs. And that with a considerable margin of other frameworks always taking at least 50\% longer (Gemini on flickr) up to Galois Pull needing 18$\times$ longer on wikipedia.  
On both synthetic graphs however, Giraph is the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.

@@

>>>>

\subsection{Single-source Shortest-paths}
For the results of our SSSP runs, we first analyze the single-node and distributed performances separately before comparing the two.

\subsubsection{Single-node}
Beginning with the single-node performance, \autoref{fig:singleNodeSSSP} shows the average calculation and execution times for SSSP on the different frameworks.

Because we did not see the perfomance going down when increasing thread counts on Galois, we always refer to Galois with 96 threads in this section (including the figures). 
For a more detailed analysis of Galois' behaviour with different thread counts, see \autoref{sec:galois_speedup}.

TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.
\begin{figure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_calcTime.png}
		\caption{Calculation times for SSSP on a single node}
		\label{fig:singleNodeSSSP_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_execTime.png}
		\caption{Execution times for SSSP on a single node}
		\label{fig:singleNodeSSSP_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeSSSP_overheadTimeNormalized.png}
		\caption{Overhead time normalized by the graph size in million edges}
		\label{fig:singleNodeSSSP_overheadNormalized}
	\end{subfigure}
	\caption{Average times on a single computation node, black bars represent one standard deviation in our testing.
	The runs on rMat28 for Ligra and Polymer failed and the frameworks were unable to complete the task.}
	\label{fig:singleNodeSSSP}
\end{figure}

\subsubsection{Distributed}

\begin{figure}
	\centering
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_calcTime.png}
		\caption{Calculation times for distributed SSSP}
		\label{fig:distributedSSSP_calc}
	\end{subfigure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedSSSP_execTime.png}
		\caption{Execution times for distributed SSSP}
		\label{fig:distributedSSSP_exec}
	\end{subfigure}
	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing.}
	\label{fig:distributedSSSP}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
	\label{fig:distributedSSSP_overhead}
\end{figure}

Moving on to the distributed cluster, \autoref{fig:distributedSSSP} shows the benchmark results as calculation and execution times.

Comparing the two Galois implementations, we find the calculation and execution times to be similar on smaller graphs and Push being the superior implementation for SSSP on larger data sets. Galois Pull being anywhere from just as fast to 3.5$\times$ slower on real-world data sets compared to the Push variant. The synthetic graphs are more extreme, execution times are close to 4$\times$ (rMat27) and 5$\times$ (rMat28) longer.

Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs, see \autoref{tbl:ssspexec}.
You can see Gemini being worse by at least a factor of 4 compared to Galois Push on all graphs except flickr.
Giraph's execution times in comparison to this are even worse, taking at least 7$\times$ longer than Galois Push on all graphs.

Evidently, Galois Push is the fastest algorithm in our lineup on 6 out of 7 graphs. With the exception being flickr, where Galois Push takes negligibly longer than the Pull counterpart.

\begin{table}
\renewcommand{\arraystretch}{1.3}
\centering
\caption{Distributed SSSP Execution Times Relative to Galois Push}
\begin{tabular}{crrrr}
\hline
\bf{Data Set}&Galois Push&Galois Pull&Gemini&Giraph\\\hline
flickr&1.0&0.97&1.02&17.46\\
orkut&1.0&1.28&4.12&14.38\\
wikipedia&1.0&1.88&5.26&7.11\\
twitter&1.0&3.56&9.79&8.76\\
rMat27&1.0&3.98&9.91&14.52\\
friendster&1.0&2.32&9.72&7.59\\
rMat28&1.0&5.70&11.07&16.5\\\hline
\end{tabular}
\label{tbl:ssspexec}
\end{table}

When taking a closer look at Giraph, it seems to not cope well with synthetic data sets. Analyzing the computation times in \autoref{fig:distributedSSSP_calc}, we see that it is the fastest framework on our real-world graphs. And that with a considerable margin of other frameworks always taking at least 50\% longer (lower bound here is Gemini on flickr) up to Galois Pull needing 18$\times$ more time on wikipedia.  
On both synthetic graphs however, Giraph is actually the slowest to compute. Giraph requires 12$\times$ or even 15$\times$ the computation time of Gemini on rMat27 or rMat28 respectively.

@@

\subsubsection{Single-Node vs. Distributed}
TODO
\begin{itemize}
	\item GaloisPush vs Galois CPU // Galois Pull verliert bei Dist.
	\item Giraph vs. Giraph
	\item Gemini vs Gemini
\end{itemize}

### 2020-08-29 documentation/paper/sections/testingMethods.tex

\subsection{Hardware}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.
Setup of each framework was performed according to our provided installation guides\footnote{The setup guides are available at \url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}.

\subsection{Benchmark Software}
All benchmarks were initiated by our benchmark script, that is available in our repository. Galois, Gemini and Giraph were benchmarked on both the 5-node cluster and a singe machine.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.

@@

\subsection{Data Sets}
The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}. We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 100 thousand edges up to an rMat28 with $4.2$ billion edges. All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
\begin{table}
	\centering
	\caption{Size Comparison of the Used Graphs}
	\begin{tabular}{crr}
		\hline
		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
		flickr&    		0.1&  2\\
		orkut&          3&    117\\
		wikipedia&      12&   378\\
		twitter&     	52&   1963\\
		rMat27&         63&   2147\\
		friendster&     68&   2586\\
		rMat28&         121&  4294\\
		\hline
	\end{tabular}
	\label{tbl:graphs}
\end{table}

@@

	We know that this is not the most realiable way for measuring the calculation times. 
	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time that is used for cleanup after calculation in the measurement.

@@

	\item Giraph has built in timers for the iterations (supersteps) which we summed up to extract the computation time.
\end{itemize}
Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.

@@

\subsection{Algorithms}
The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
For frameworks that support multiple implementations (e.g. PageRank in push and pull modes), we included the alternatives in our testing. 
\autoref{tbl:algorithms} shows the tested algorithms in detail.

\newcommand{\txtdag}{\textsuperscript{\textdagger}}
\begin{table}
	\renewcommand{\arraystretch}{1.3}
	\centering
	\caption{Tested Algorithms of Each Framework}
	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
		PageRank&$\Delta$Push, $\Delta$Pull&&&&Regular, $\Delta$\\\hline
		SSSP&?, Push, Pull&&&&\\\hline
		BFS&Yes&&Yes*&&\\
		\hline
	\end{tabularx}
	\label{tbl:algorithms}

	{\vspace*{1ex}* Algorithm not natively supported

	\txtdag{} Modified algorithm

	\textsuperscript{d} Single node and distributed
	}

\end{table}

>>>>

\section{Testing Methods}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.

Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
All benchmark cases were initiated by our benchmark script available in our repository. 
All five frameworks are tested on a single server. 
Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.

@@

\subsection{Data Sets}
The graphs used in our testing can be seen in detail in \autoref{tbl:graphs}. We included a variaty of different graph sizes, from relatively small graphs like the flickr graph with 2 million edges up to an rMat28 with 4.2 billion edges. All graphs except the rMat27 and rMat28 are exemplary real-world graphs and were retrieved from the graph database\footnote{\url{http://konect.uni-koblenz.de/}} associated with the Koblenz Network Collection (KONECT)\cite{konect}.
Both the rMat27 and rMat28 were created with a modified version (we changed the output format to EdgeList) of a graph generator provided by Ligra.
\begin{table}
	\centering
	\caption{Size Comparison of the Used Graphs}
	\begin{tabular}{crr}
		\hline
		\bf{Graph}&\# Vertices (M)&\# Edges (M)\\\hline
		flickr&    		0.1&  2\\
		orkut&          3&    117\\
		wikipedia&      12&   378\\
		twitter&     	52&   1963\\
		rMat27&         63&   2147\\
		friendster&     68&   2586\\
		rMat28&         121&  4294\\
		\hline
	\end{tabular}
	\label{tbl:graphs}
\end{table}

@@

	We know that this is not the most realiable way for measuring the calculation times. 
	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time in the measurement that is used for cleanup after calculation.

@@

	\item Giraph has built in timers for the iterations (supersteps), the sum of those is the computation time.
\end{itemize}
Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.

@@

\subsection{Algorithms}
The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
We always show the results of PageRank with a maximum of five iterations.
For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included the alternatives in our testing.

In detail, the algorithms for each framework are:
\begin{itemize}
	\item Galois: CPU, Push, Pull für PR und SSSP..
	\item Gemini:
	\item Giraph: Giraph does not natively supply a BFS algorithm, so in our comparisons a custom implementation is used.
	\item Ligra:
	\item Polymer: Polymer supports two implementations of Pagerank
\end{itemize}

% \autoref{tbl:algorithms} shows the tested algorithms in detail.

% \newcommand{\txtdag}{\textsuperscript{\textdagger}}
% \newcommand{\singleNode}{\textsuperscript{s}}
% \newcommand{\distributed}{\textsuperscript{d}}
% \begin{table}
% 	\renewcommand{\arraystretch}{1.3}
% 	\centering
% 	\caption{REMOVE THIS // Tested Algorithms of Each Framework}
% 	\begin{tabularx}{0.9\columnwidth}{cYYYYY}
% 		\bf{Algorithm}&Galois&Gemini&Giraph&Ligra&Polymer\\\hline
% 		PageRank&Yes\singleNode{}, $\Delta$Push\distributed{}, $\Delta$Pull\distributed{}&&Yes\singleNode\distributed&&Regular\singleNode, $\Delta$\singleNode\\\hline
% 		SSSP&?, Push, Pull&&Yes\singleNode\distributed\txtdag&&Yes\singleNode{}\\\hline
% 		BFS&Yes&&Yes\singleNode\distributed{}*&&\\
% 		\hline
% 		\multicolumn{6}{p{0.5\columnwidth}}{* Algorithm not natively supported

% 		\txtdag{} Modified algorithm

% 		\distributed{} distributed

% 		\singleNode{} single-node}
% 	\end{tabularx}
% 	\label{tbl:algorithms}

% \end{table}
